---
title: "Analysis"
author: "Tim Folsom"
date: "7/31/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

This vignette will cover two typical analyses that HEI scores might be used in and will highlight how this package is employed in the process.

```{r, echo=TRUE}
library(hei)
library(ggplot2)
library(dplyr)
```

We'll be using data for both days (which is the default).

```{r, echo=TRUE, results='hide'}

fped0910 <- get_fped("2009/2010", day = "both")

diet0910 <- get_diet("2009/2010", day = "both")

demo0910 <- get_demo("2009/2010")

hei0910 <- hei(fped0910,diet0910,demo0910)
```

## Analysis 1: BMI

We will now conduct a linear regression analysis to examine the relationship between Body Mass Index and HEI scores. First select only the relevant columns from the output data set.

```{r, echo=TRUE}
hei0910 <- hei0910 %>% 
    select(SEQN, RIDAGEYR, HEI)
```

Then, using the nhanesA package, pull down NHANES body measures data from the web. Select the relevant columns and filter out those rows that contain missing data.

```{r, echo=TRUE, results='hide'}
BMX_0910 <- nhanesA::nhanes('BMX_F') %>%
    select(SEQN, BMXBMI) %>% 
    filter(!is.na(BMXBMI))
```

Merge the BMI data with our HEI data. We will also restrict our analysis to adults (age 20 and over).

```{r, echo=TRUE}
heibmi0910 <- merge(hei0910, BMX_0910, by = "SEQN") %>%
    filter(RIDAGEYR > 19)
```

We can produce a scatter plot (using a function from the car package) visualizing the distributions of BMI and HEI scores as well as their relationship to each other.

```{r, echo=TRUE, message=TRUE}
ggplot(heibmi0910, aes(HEI, BMXBMI)) +
    geom_point(alpha = 0.25) +
    geom_smooth(method = "lm")
```

The distribution of HEI scores appears to be normally distributed (an assumption of linear regression).

```{r, echo=TRUE, message=TRUE}
hist(heibmi0910$HEI)
```

We see that HEI is a highly significant predictor of BMI, though the effect size is small.

```{r, echo=TRUE, message=TRUE}
heibmi0910.lm <- lm(BMXBMI ~ HEI, data=heibmi0910)
summary(heibmi0910.lm)
```


## Analysis 2: Food Security

This next example will analyze the relationship between HEI and food security. As before, we first select only the relevant columns from the original output data set.

```{r, echo=TRUE}
hei0910 <- hei0910 %>% 
    select(SEQN, RIDAGEYR, HEI)
```

Then, using the nhanesA package, pull down NHANES food security data from the web. Select the relevant columns and filter out those rows that contain missing data.

```{r, echo=TRUE, results='hide'}
FSQ_0910 <- nhanesA::nhanes('FSQ_F') %>%
    select(SEQN, FSDAD) %>% 
    filter(!is.na(FSDAD))
```

Merge the food security data with our HEI data. We will also restrict our analysis to adults (age 20 and over).

```{r, echo=TRUE}
heifsq0910 <- merge(hei0910, FSQ_0910, by = "SEQN") %>%
    filter(RIDAGEYR > 19)
```

We can produce a bar plot to approximate the distribution of individuals across the four categories (1 meaning totally secure, 4 representing very little security).

```{r, echo=TRUE}
ggplot(heifsq0910, aes(FSDAD)) + geom_bar()
```

The average score for those in group 1 appears to be quite a bit higher than the rest of the groups.

```{r, echo=TRUE, message=TRUE}
heifsq0910 %>% 
    group_by(FSDAD) %>% 
    summarise(mean(HEI))
```

We see that the difference between the secure group and each of the other groups is highly significant.

```{r, echo=TRUE, message=TRUE}
heifsq0910$FSDAD <- relevel(factor(heifsq0910$FSDAD), ref="1")
heifsq0910.lm <- lm(HEI ~ FSDAD, data=heifsq0910)
summary(heifsq0910.lm)
```